---
title: "Alcohol_Cannabis_code"
author: Zhaoying YU
output: html_document
date: "2023-10-16"
---
## Set-up

```{r, purl=FALSE}
library("dplyr")
library("tidyverse")
```

## Import data
```{r, purl=FALSE}
dat <-
  readRDS("/Users/lilian/Desktop/dat_Anonymous.rds")
```

## Alcohol algorithms
# Adjust AUDIT score
adjusting item scores to 0-4 as indicated: 
Frequency (AUDIT1) > f.29091-0.0
scored 0-4, no adjustment needed 

```{r}
if(diagnosticsFlag) unique(dat$`29091-0.0`)
dat <- dat %>%
  mutate(
    `29091-0.0` = recode(
      as.character(`29091-0.0`),
      "0" = 0,
      "1" = 1,
      "2" = 2,
      "3" = 3,
      "4" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```


Typical drinks (AUDIT 1a) > f.29092-0.0
scored 1-5, needs adjusting 

```{r}
if(diagnosticsFlag) unique(dat$`29092-0.0`)
dat <- dat %>%
  mutate(
    `29092-0.0` = recode(
      as.character(`29092-0.0`),
      "1" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Six or more drinks (AUDIT 1b) > f.29093-0.0
scored 1-5, needs adjusting 
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29093-0.0`)
dat <- dat %>%
  mutate(
    `29093-0.0` = recode(
      as.character(`29093-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Unable to stop (AUDIT 2) > f.29094-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29094-0.0`)
dat <- dat %>%
  mutate(
    `29094-0.0` = recode(
      as.character(`29094-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```


Failed to do what expected due to drinking (AUDIT 3) > f.29095-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29095-0.0`)
dat <- dat %>%
  mutate(
    `29095-0.0` = recode(
      as.character(`29095-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Needed to drink first thing (AUDIT 4) > f.29096-0.0
scored 1-5, needs adjusting 
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29096-0.0`)
dat <- dat %>%
  mutate(
    `29096-0.0` = recode(
      as.character(`29096-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Guilt due to drinking (AUDIT 5) f.29097-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29097-0.0`)
dat <- dat %>%
  mutate(
    `29097-0.0` = recode(
      as.character(`29097-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Unable to remember due to drink (AUDIT 6) f.29098-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29098-0.0`)
dat <- dat %>%
  mutate(
    `29098-0.0` = recode(
      as.character(`29098-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Injury due to drinking ever (AUDIT 7) f.29099-0.0
scored 0,1,2, needs adjusting to 0,2,4 

```{r}
if(diagnosticsFlag) unique(dat$`29099-0.0`)
dat <- dat %>%
  mutate(
    `29099-0.0` = recode(
      as.character(`29099-0.0`),
      "0" = 0,
      "1" = 2,
      "2" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Advice to cut down ever (AUDIT 8) f.29100-0.0
scored 0,1,2, needs adjusting to 0,2,4

```{r}
if(diagnosticsFlag) unique(dat$`29100-0.0`)
dat <- dat %>%
  mutate(
    `29100-0.0` = recode(
      as.character(`29100-0.0`),
      "0" = 0,
      "1" = 2,
      "2" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

# Sum of AUDIT score

```{r}
dat[c(
  "29091-0.0",
  "29092-0.0",
  "29093-0.0",
  "29094-0.0",
  "29095-0.0",
  "29096-0.0",
  "29097-0.0",
  "29098-0.0",
  "29099-0.0",
  "29100-0.0"
)] <- lapply(
  dat[c(
    "29091-0.0",
    "29092-0.0",
    "29093-0.0",
    "29094-0.0",
    "29095-0.0",
    "29096-0.0",
    "29097-0.0",
    "29098-0.0",
    "29099-0.0",
    "29100-0.0"
  )],
  as.numeric
)

#Rewrites Sum_AUDIT to NA if all AUDIT scores are missing/NA
dat$Sum_AUDIT <- apply(
  dat[, c(
    "29091-0.0",
    "29092-0.0",
    "29093-0.0",
    "29094-0.0",
    "29095-0.0",
    "29096-0.0",
    "29097-0.0",
    "29098-0.0",
    "29099-0.0",
    "29100-0.0"
  )],
  1,
  function(row) {
    if (all(
      is.na(row)
    )
    ) {
      NA
    } else {
      sum(row,
        na.rm = TRUE
      )
    }
  }
)
```

# Hazardous / harmful alcohol use (12 month) 
Score 8-15 predicts “hazardous drinking” while scores of 16 and over predict “harmful drinking” 
Hazardous alcohol use: Score {AUDIT full scale} >= 8 & <= 15 (when items scored 0-4) 
Harmful alcohol use: Score {AUDIT full scale} >= 16 (when items scored 0-4) 

```{r}
#Hazardous use
dat <- dat %>%
  mutate(
    alcohol_hazardous_use = case_when(
      is.na(Sum_AUDIT) ~ NA_real_,
      Sum_AUDIT >= 8 ~ 1,
      TRUE ~ 0
    )
  )

#Harmful use
dat <- dat %>%
  mutate(
    alcohol_harmful_use = case_when(
      is.na(Sum_AUDIT) ~ NA_real_,
      Sum_AUDIT >= 16 ~ 1,
      TRUE ~ 0
    )
  )
```

# Non-case for hazardous / harmful alcohol use
AUDIT full scale score under cut-off (8), excluding those who reported alcohol addiction in this questionnaire or reported at baseline they had stopped drinking due to illness, on drs advice or as a health precaution

{AUDIT score} <8

AND 
NOT {ever alcohol dependence} 
>Used data field 20117: alcohol drinker status = never. (-3) Prefer not to answer, (0) Never, (1) Previous, (2) Current
> '20117-0.0' == 0 

AND 
NOT reason for reducing amount of alcohol drunk 2664 = “ill health”, “doctor’s advice”or “health precaution” [1,2or3]
>!('2664-0.0' %in% c(1,2,3))

+++JZ: We do not have 20117-0.0 (20117: alcohol drinker status) in our test data.
The algorithm requests “ever alcohol dependence”. Should this be coded using any addiction diagnoses or the addiction fields? https://biobank.ndph.ox.ac.uk/showcase/label.cgi?id=141
Vast majority of UKB have 20117-0.0==1.


```{r}
dat <- dat %>%
  mutate(
    alcohol_nonuse = case_when(
      Sum_AUDIT < 8 &
        `20117-0.0` == 0 &
        `2664-0.0` %in% c(4, 5)
      ~ 1,
      Sum_AUDIT >= 8 |
        `20117-0.0` %in% c(1, 2) |
        `2664-0.0` %in% c(1, 2, 3)
      ~ 0,
      TRUE ~ NA_real_
    )
  )
```


## Cannabis algorithms
# Cannabis use ever
Used cannabis (CU1) = Yes, 1-2 times (01) OR Yes, 3-10 times (02) OR Yes, 11-100 times (03) Yes, more than 100 times (04)
>'29104-0.0' %in% c(1,2,3,4)

```{r}
dat <- dat %>%
  mutate(
    cannabis_use = case_when(
      `29104-0.0` %in% c(1, 2, 3, 4) ~ 1,
      `29104-0.0` == 0 ~ 0,
      TRUE ~ NA_integer_
    )
  )
```

# Daily cannabis use ever 
Maximum frequency of taking cannabis when using is every day
How often = Every day (04) 
>Every day is not 04 as suggested in algorithm, but is 3 as shown on UKbiobank showcase here:https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=1952
>'29107-0.0' == 3

```{r}
if(diagnosticsFlag) unique(dat$`29107-0.0`)
dat <- dat %>%
  mutate(
    daily_cannabis_use = case_when(
      `29107-0.0` == 3 ~ 1,
      `29107-0.0` >= 0 & `29107-0.0` < 3 ~ 0,
      TRUE ~ NA_integer_
    )
  )
```
