---
title: "Alcohol_Cannabis_code"
author: Zhaoying YU
output: html_document
date: "2023-10-16"
---
## Set-up

```{r, purl=FALSE}
library("dplyr")
library("tidyverse")
```

## Import data
```{r, purl=FALSE}
dat <-
  readRDS("/Users/lilian/Desktop/dat_Anonymous.rds")
```

## Alcohol algorithms
# Adjust AUDIT score
adjusting item scores to 0-4 as indicated: 
Frequency (AUDIT1) > f.29091-0.0
scored 0-4, no adjustment needed 

```{r}
if(diagnosticsFlag) unique(dat$`29091-0.0`)
dat <- dat %>%
  mutate(
    `29091-0.0` = recode(
      as.character(`29091-0.0`),
      "0" = 0,
      "1" = 1,
      "2" = 2,
      "3" = 3,
      "4" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```


Typical drinks (AUDIT 1a) > f.29092-0.0
scored 1-5, needs adjusting 

```{r}
if(diagnosticsFlag) unique(dat$`29092-0.0`)
dat <- dat %>%
  mutate(
    `29092-0.0` = recode(
      as.character(`29092-0.0`),
      "1" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Six or more drinks (AUDIT 1b) > f.29093-0.0
scored 1-5, needs adjusting 
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29093-0.0`)
dat <- dat %>%
  mutate(
    `29093-0.0` = recode(
      as.character(`29093-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Unable to stop (AUDIT 2) > f.29094-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29094-0.0`)
dat <- dat %>%
  mutate(
    `29094-0.0` = recode(
      as.character(`29094-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```


Failed to do what expected due to drinking (AUDIT 3) > f.29095-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29095-0.0`)
dat <- dat %>%
  mutate(
    `29095-0.0` = recode(
      as.character(`29095-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Needed to drink first thing (AUDIT 4) > f.29096-0.0
scored 1-5, needs adjusting 
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29096-0.0`)
dat <- dat %>%
  mutate(
    `29096-0.0` = recode(
      as.character(`29096-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Guilt due to drinking (AUDIT 5) f.29097-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29097-0.0`)
dat <- dat %>%
  mutate(
    `29097-0.0` = recode(
      as.character(`29097-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Unable to remember due to drink (AUDIT 6) f.29098-0.0
scored 1-5, needs adjusting
>scores are 0, 2, 3, 4, 5. Adjusted to be 0, 1, 2, 3, 4

```{r}
if(diagnosticsFlag) unique(dat$`29098-0.0`)
dat <- dat %>%
  mutate(
    `29098-0.0` = recode(
      as.character(`29098-0.0`),
      "0" = 0,
      "2" = 1,
      "3" = 2,
      "4" = 3,
      "5" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Injury due to drinking ever (AUDIT 7) f.29099-0.0
scored 0,1,2, needs adjusting to 0,2,4 

```{r}
if(diagnosticsFlag) unique(dat$`29099-0.0`)
dat <- dat %>%
  mutate(
    `29099-0.0` = recode(
      as.character(`29099-0.0`),
      "0" = 0,
      "1" = 2,
      "2" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

Advice to cut down ever (AUDIT 8) f.29100-0.0
scored 0,1,2, needs adjusting to 0,2,4

```{r}
if(diagnosticsFlag) unique(dat$`29100-0.0`)
dat <- dat %>%
  mutate(
    `29100-0.0` = recode(
      as.character(`29100-0.0`),
      "0" = 0,
      "1" = 2,
      "2" = 4,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )
```

# Sum of AUDIT score

```{r}
dat[c(
  "29091-0.0",
  "29092-0.0",
  "29093-0.0",
  "29094-0.0",
  "29095-0.0",
  "29096-0.0",
  "29097-0.0",
  "29098-0.0",
  "29099-0.0",
  "29100-0.0"
)] <- lapply(
  dat[c(
    "29091-0.0",
    "29092-0.0",
    "29093-0.0",
    "29094-0.0",
    "29095-0.0",
    "29096-0.0",
    "29097-0.0",
    "29098-0.0",
    "29099-0.0",
    "29100-0.0"
  )],
  as.numeric
)

#Rewrites MHQ2.AUDITScore to NA if all AUDIT scores are missing/NA
dat$MHQ2.AUDITScore <- apply(
  dat[, c(
    "29091-0.0",
    "29092-0.0",
    "29093-0.0",
    "29094-0.0",
    "29095-0.0",
    "29096-0.0",
    "29097-0.0",
    "29098-0.0",
    "29099-0.0",
    "29100-0.0"
  )],
  1,
  function(row) {
    if (all(
      is.na(row)
    )
    ) {
      NA
    } else {
      sum(row,
        na.rm = TRUE
      )
    }
  }
)
```

# Hazardous / harmful alcohol use (12 month) 
Score 8-15 predicts “hazardous drinking” while scores of 16 and over predict “harmful drinking” 
Hazardous alcohol use: Score {AUDIT full scale} >= 8 & <= 15 (when items scored 0-4) 
Harmful alcohol use: Score {AUDIT full scale} >= 16 (when items scored 0-4) 

```{r}
#Hazardous use
dat <- dat %>%
  mutate(
    MHQ2.AlcoholHazardousHarmfulUseCase = case_when(
      is.na(MHQ2.AUDITScore) ~ NA_real_,
      MHQ2.AUDITScore >= 8 ~ 1,
      TRUE ~ 0
    )
  )

#Harmful use
dat <- dat %>%
  mutate(
    MHQ2.AlcoholHarmfulDrinking = case_when(
      is.na(MHQ2.AUDITScore) ~ NA_real_,
      MHQ2.AUDITScore >= 16 ~ 1,
      TRUE ~ 0
    )
  )
```

## Cannabis algorithms
# Cannabis use ever
Used cannabis (CU1) = Yes, 1-2 times (01) OR Yes, 3-10 times (02) OR Yes, 11-100 times (03) Yes, more than 100 times (04)
>'29104-0.0' %in% c(1,2,3,4)

```{r}
dat <- dat %>%
  mutate(
    MHQ2.CannabisUseEver = case_when(
      `29104-0.0` %in% c(1, 2, 3, 4) ~ 1,
      `29104-0.0` == 0 ~ 0,
      TRUE ~ NA_integer_
    )
  )
```

# Daily cannabis use ever 
Maximum frequency of taking cannabis when using is every day
How often = Every day (04) 
>Every day is not 04 as suggested in algorithm, but is 3 as shown on UKbiobank showcase here:https://biobank.ndph.ox.ac.uk/showcase/coding.cgi?id=1952
>'29107-0.0' == 3

```{r}
if(diagnosticsFlag) unique(dat$`29107-0.0`)
dat <- dat %>%
  mutate(
    MHQ2.CannabisUseDaily = case_when(
      `29107-0.0` == 3 ~ 1,
      `29107-0.0` >= 0 & `29107-0.0` < 3 ~ 0,
      TRUE ~ NA_integer_
    )
  )
```
