---
title: "MHQ2 Anorexia nervosa DSM-5 algoritm"
author: "Zhaoying Yu, Helena Davies, Christopher Huebel, Johan Z"
output: html_document
date: "2023-10-07"
---

```{r}
library("tidyverse")
```


# Recode multiple choice question
29130 Methods of controlling body shape or weight when at this low weight
1	Made yourself vomit
2	Used laxatives (pills or liquids)
3	Used diuretics (water pills)
4	Used weight loss pills
5	Exercised excessively or felt distressed if unable to exercise
6	Fasted or not eaten for eight waking hours or more
7	Other methods to lose weight/stay at low weight
0	None of above
-3	Prefer not to answer

```{r made yourself vomit}
dat <- dat %>%
  mutate(
    LW_made_yourself_vomit =
      case_when(
        `29130-0.0` == 1 |
        `29130-0.1` == 1 |
        `29130-0.2` == 1 | 
        `29130-0.3` == 1 |
        `29130-0.4` == 1 |
        `29130-0.5` == 1 |
        `29130-0.6` == 1 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 1.021)
dat %>%
  count(LW_made_yourself_vomit)
```

```{r used laxatives}
dat <- dat %>%
  mutate(
    LW_used_laxatives =
      case_when(
        `29130-0.0` == 2 |
        `29130-0.1` == 2 |
        `29130-0.2` == 2 | 
        `29130-0.3` == 2 |
        `29130-0.4` == 2 |
        `29130-0.5` == 2 |
        `29130-0.6` == 2 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 1.050)
dat %>%
  count(LW_used_laxatives)
```

```{r used diuretics}
dat <- dat %>%
  mutate(
    LW_used_diuretics =
      case_when(
        `29130-0.0` == 3 |
        `29130-0.1` == 3 |
        `29130-0.2` == 3 | 
        `29130-0.3` == 3 |
        `29130-0.4` == 3 |
        `29130-0.5` == 3 |
        `29130-0.6` == 3 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 247)
dat %>%
  count(LW_used_diuretics)
```

```{r used weight loss pills}
dat <- dat %>%
  mutate(
    LW_used_weight_loss_pills =
      case_when(
        `29130-0.0` == 4 |
        `29130-0.1` == 4 |
        `29130-0.2` == 4 | 
        `29130-0.3` == 4 |
        `29130-0.4` == 4 |
        `29130-0.5` == 4 |
        `29130-0.6` == 4 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 288)
dat %>%
  count(LW_used_weight_loss_pills)
```

```{r excessive exercise}
dat <- dat %>%
  mutate(
    LW_excessive_exercise =
      case_when(
        `29130-0.0` == 5 |
        `29130-0.1` == 5 |
        `29130-0.2` == 5 | 
        `29130-0.3` == 5 |
        `29130-0.4` == 5 |
        `29130-0.5` == 5 |
        `29130-0.6` == 5 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 1,917)
dat %>%
  count(LW_excessive_exercise)
```

```{r fasted}
dat <- dat %>%
  mutate(
    LW_fasted =
      case_when(
        `29130-0.0` == 6 |
        `29130-0.1` == 6 |
        `29130-0.2` == 6 | 
        `29130-0.3` == 6 |
        `29130-0.4` == 6 |
        `29130-0.5` == 6 |
        `29130-0.6` == 6 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 2,403)
dat %>%
  count(LW_fasted)
```

```{r other weight control}
dat <- dat %>%
  mutate(
    LW_other_weight_control =
      case_when(
        `29130-0.0` == 7 |
        `29130-0.1` == 7 |
        `29130-0.2` == 7 | 
        `29130-0.3` == 7 |
        `29130-0.4` == 7 |
        `29130-0.5` == 7 |
        `29130-0.6` == 7 ~ 1,
                     
        # did not engage in any weight control behaviour
        `29130-0.0` == 0 ~ 0,
        
        # people who have not answered the MHQ2
        TRUE ~ NA_real_
                     )
         )

# Check (expecting 823)
dat %>%
  count(LW_other_weight_control)
```

# BMI at low weight

Lowest weight during period when underweight [metric] / Height [cm]/100)^2

```{r}
dat <- dat %>%
  mutate(
    BMI_at_low_weight =
      case_when(
        is.na(`29125-0.0`) |
          is.na(`50-0.0`) |
          `29125-0.0` %in% c(0, -1, -3) ~ NA_real_,
        TRUE ~ `29125-0.0` / (`50-0.0`/100)^2)
        )
```


```{r}
dat %>% 
  select(
    BMI_at_low_weight
  ) %>% 
  summary()
```


# Anorexia nervosa

```{r}
dat <- dat %>%
  mutate(Anorexia_nervosa = 
           case_when(
# Criterion A             
           # BMI at low weight <= 18.55 kg/m2
                       BMI_at_low_weight <= 18.55 &
# Criterion B
             # Afraid that you might gain weight or become fat = 1 Yes
                       (`29123-0.0` == 1 |

           # Have you done any of the following as a way to control your body shape or weight? = any of 1 to 7
             # when at low weight
                        LW_made_yourself_vomit == 1 |
                        LW_used_laxatives == 1 |
                        LW_used_diuretics == 1 |
                        LW_used_weight_loss_pills == 1 |
                        LW_excessive_exercise == 1 |
                        LW_fasted == 1 |
                        LW_other_weight_control == 1) &
# Criterion C  
           # Did you feel fat? = 1 Yes
                       (`29122-0.0` == 1 |

           # During the time when you were at this low weight: Did you think or feel that your body or parts of your body were larger than they actually were? = 1 Yes
                        `29124-0.0` == 1 |

           # Did/do you ever think your low weight had/has negative consequences for your health? = 0 No
                        `29128-0.0` == 0 |

           # Was your self-esteem dependent on your body shape or weight = 2 A great deal OR 1 A moderate amount
                        `29129-0.0` %in% c(1, 2) )
            ~ 1,
           
           # has answered ED section of MHQ2
           !is.na(`29205-0.0`) ~ 0,
           
           # people who have not answered the MHQ2
           TRUE ~ NA_real_
         )
  )
```


```{r}
dat %>%
  count(Anorexia_nervosa)
```


# Anorexia nervosa - binge-eating/purging subtype

```{r}
dat <- dat %>%
  mutate(Anorexia_nervosa_binge_eating_purging =
# Criterion A
           # BMI at low weight <= 18.55 kg/m2
              case_when( BMI_at_low_weight <= 18.55 &
# Criterion B
           # Afraid that you might gain weight or become fat = 1 Yes
                       (`29123-0.0` == 1 |

          # Have you done any of the following as a way to control your body shape or weight? = any of 1 to 7
            # when at low weight
                        made_yourself_vomit == 1 |
                        used_laxatives == 1 |
                        used_diuretics == 1 |
                        used_weight_loss_pills == 1 |
                        excessive_exercise == 1 |
                        fasted == 1 |
                        other_weight_control == 1) &

# Criterion C
           # Did you feel fat? = 1 Yes
                       (`29122-0.0` == 1 |

           # During the time when you were at this low weight: Did you think or feel that your body or parts of your body were larger than they actually were? = 1 Yes
                        `29124-0.0` == 1 |

           # Did/do you ever think your low weight had/has negative consequences for your health? = 0 No
                        `29128-0.0` == 0 |

           # Was your self-esteem dependent on your body shape or weight = 2 A great deal OR 1 A moderate amount
                        `29129-0.0` %in% c(1,2) ) &

# AN binge-eating criterion
          # Have you had recurrent episodes of excessive overeating or binge eating (EP6c) = 2 Yes, ONLY at time(s) of low weight OR 1 Yes, BOTH at time(s) of low weight AND at time(s) when I was not at low weight)
                       (`29134-0.0` %in% c(1,2) |
# AN purging criterion
          # Have you done any of the following as a way to control your body shape or weight (EP5) = 1 Made yourself vomit OR 2 Used laxatives (pills or liquids meant to stimulate bowel movement) OR 3 Used diuretics (water pills) 
            # when at low weight
                        LW_made_yourself_vomit == 1 |
                        LW_used_laxatives == 1 |
                        LW_used_diuretics == 1)
            ~ 1,
           
           # has answered ED section of MHQ2
           !is.na(`29205-0.0`) ~ 0,
           
           # people who have not answered the MHQ2
           TRUE ~ NA_real_
         )
  )
```


```{r}
dat %>%
  count(Anorexia_nervosa_binge_eating_purging)
```

# Anorexia nervosa - restricting subtype

```{r}
dat <- dat %>%
  mutate(Anorexia_nervosa_restricting =
# Criterion A
           # BMI at low weight <= 18.55 kg/m2
              case_when( BMI_at_low_weight <= 18.55 &
# Criterion B
           # Afraid that you might gain weight or become fat = 1 Yes
                       (`29123-0.0` == 1 |

          # Have you done any of the following as a way to control your body shape or weight? = 5 exercise excessively OR 6 fasted
                        LW_excessive_exercise == 1 |
                        LW_fasted == 1 |
                        LW_used_weight_loss_pills == 1) &

# Criterion C
           # Did you feel fat? = 1 Yes
                       (`29122-0.0` == 1 |

           # During the time when you were at this low weight: Did you think or feel that your body or parts of your body were larger than they actually were? = 1 Yes
                        `29124-0.0` == 1 |

           # Did/do you ever think your low weight had/has negative consequences for your health? = 0 No
                        `29128-0.0` == 0 |

           # Was your self-esteem dependent on your body shape or weight = 2 A great deal OR 1 A moderate amount
                        `29129-0.0` %in% c(1,2) ) &

# Restricting type: During the last 3 months... 
  
  #...the individual has not engaged in recurrent episodes of binge eating or purging behavior (i.e., self-induced vomiting or the misuse of laxatives, diuretics, or enemas)...
                        (`29132-0.0` == 0 | # No
                        `29134-0.0` == 2 # Only at time(s) when I was not at low weight [2,767]
                        ) &
  # During the period of time when you were at your lowest weight, did you use any of the following as a way to control your weight or shape?
                      (
                        LW_made_yourself_vomit == 0 | # No
                        LW_used_laxatives == 0 | # No
                        LW_used_diuretics == 0 # No
                        )

            ~ 1,
           
           # has answered ED section of MHQ2
           !is.na(`29205-0.0`) ~ 0,
           
           # people who have not answered the MHQ2
           TRUE ~ NA_real_
         )
  )
```



```{r}
dat %>%
  count(Anorexia_nervosa_restricting)
```

# Check overlap between ANR and ANBP

```{r}
dat %>%
  group_by(Anorexia_nervosa_restricting, Anorexia_nervosa_binge_eating_purging) %>%
  tally()
```



```{r}
dat %>%
  select(
    Anorexia_nervosa,
    Anorexia_nervosa_binge_eating_purging,
    Anorexia_nervosa_restricting,
  ) %>%
  filter(
    Anorexia_nervosa == 1
  )
```

