---
title: "Anxiety"
author: "Abigail ter Kuile"
date: "2023-10-06"
output: html_document
---

# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  prompt = FALSE,
  cache = FALSE
)
```

Install packages
```{r Install packages}
#install.packages("tidyverse")
```

Load packages from library 
```{r Load packages from library}
library(summarytools)
library(tidyverse)
```

```{r Clear global environment}
rm(list=ls())
```

```{r Get system date}
date <- Sys.Date()
date
```

## Read in data
/Users/abiterkuile/OneDrive/UCL/Postdoc/ukb/mhq2/anon_data/MHQ2_Anonymous.rds
```{r read in pheno_covar file}

MHQ <- readRDS(file = "/Users/abiterkuile/OneDrive/UCL/Postdoc/ukb/mhq2/anon_data/MHQ2_Anonymous.rds")
               
```

# Make anxiety & panic field data frame (for easier inspecting)
Anxiety & panic field names MHQ: 29058-29075
https://biobank.ndph.ox.ac.uk/ukb/label.cgi?id=1504

# specify field numbers
```{r}
anx_pad_field_numbers <- as.character(29058:29075)
```

# select anxiety & panic field numbers into dataframe
```{r}
anx_pad_data <- MHQ %>% 
  select(matches(paste(anx_pad_field_numbers, 
                       collapse = "|")))

# Inspect data
glimpse(anx_pad_data)
summary(anx_pad_data)
```
# Convert data to numeric and recode -3 (prefer not to answer) to NA
To avoid overwriting raw data, save changes in new columns with suffix "_mod" (modified)
```{r}
anx_pad_data <- anx_pad_data %>%
  mutate(across(everything(), ~ ., .names = "{.col}_mod")) %>%
  mutate_all(~ as.numeric(.)) %>%
  mutate_all(~ case_when(
    . == -3 ~ NA_real_,
    TRUE ~ .
  ))
```

# Inspect data of modified columns
```{r}
modified_columns <- anx_pad_data %>%
  select(ends_with("_mod"))

# Inspect modified columns
glimpse(modified_columns)
summary(modified_columns)
```


# Generalised Anxiety Disorder 

## GAD-7 - Full Score
Sum scores on [0-3] individual items {Feeling nervous, anxious or on edge (29058), Not being able to stop or control worrying (29059), Worrying too much about different things (29060), Trouble relaxing (29061), Being so restless that it is hard to sit still (29062), Becoming easily annoyed or irritable (29063), Feeling afraid as if something awful might happen (29064)} 

https://biobank.ndph.ox.ac.uk/ukb/coding.cgi?id=168
0	Not at all
1	Several days
2	More than half the days
3	Nearly every day
-3	Prefer not to answer 

# Sum GAD-7 scores
Check response minimum is 0 and max is 21
```{r}
anx_pad_data <- anx_pad_data %>%
  mutate(
    Gad7Score = rowSums(select(., 
                               "29058-0.0_mod",
                               "29059-0.0_mod",
                               "29060-0.0_mod",
                               "29061-0.0_mod",
                               "29062-0.0_mod",
                               "29063-0.0_mod",
                               "29064-0.0_mod"),
                        na.rm = TRUE))

summary(anx_pad_data$Gad7Score)

anx_pad_data %>%
  freq(Gad7Score)
```
Create column that counts the number of NAs for GAD7 score
```{r}
anx_pad_data <- anx_pad_data %>%
  mutate(
    Gad7NaCount = rowSums(select(., 
      "29058-0.0_mod",
      "29059-0.0_mod",
      "29060-0.0_mod",
      "29061-0.0_mod",
      "29062-0.0_mod",
      "29063-0.0_mod",
      "29064-0.0_mod"
    ) %>% is.na())
  )

# check NA count
anx_pad_data %>%
  select(Gad7NaCount,
         "29058-0.0_mod",
      "29059-0.0_mod",
      "29060-0.0_mod",
      "29061-0.0_mod",
      "29062-0.0_mod",
      "29063-0.0_mod",
      "29064-0.0_mod") %>% 
  head()

#check freq
anx_pad_data %>%
  freq(Gad7NaCount)
```

## Generalised Anxiety Disorder Current (GAD-7 derived GAD case status)

Note: This is not consistent with MHQ1 {current anxiety}, where CIDI-SF(L) GAD module was used, and participants were only categorised as current anxiety if positive on the CIDI GAD ever. The CIDI-SF GAD is not included in the MHQ2.

GAD7 full score >= 10. (10 is moderate anxiety). Ref: Kroenke K, Spitzer RL, Williams JB, Löwe B. The patient health questionnaire somatic, anxiety, and depressive symptom scales: a systematic review. Gen Hosp Psychiatry. 2010;32(4):345-59

+++ ATK: note that there is no control algorithm criteria in the MHQ2 algorithm guidance. Confirm with team whether or not the below handling of controls should be added. We have added the following criteria (not specified in MHQ2 algorithm doc): 0 is assigned if they score less than 10, and in case they could score on the missed NA items [max 3 each] taking them above a threshold of 10, this is handled by including a threshold where their score + NA count x3 does not take them above 10
```{r}
# Create a new column named 'GadCurrent' in the data frame and calculate its value.
anx_pad_data <- anx_pad_data %>%
  mutate(
    GadCurrent = case_when(
      is.na(Gad7Score) ~ NA_real_, # If 'Gad7Score' is missing (NA) Assign NA to the 'GadCurrent' column.
      Gad7Score > 9 ~ 1, # If 'Gad7Score' is greater than 9 (i.e. >=10 for moderate current Gad) Assign 1 to the 'GadCurrent' column (as case)
        Gad7Score < 10 & Gad7Score + Gad7NaCount*3 < 10 ~ 0, # if score less than 10, and adding score with NA count *3 (in case they score on the missed NA items [max 3 each] taking them above threshold of 10) - Assign 0 
        TRUE ~ NA_real_ # NA if no conditions are met
        )
      )
    
anx_pad_data %>%
  freq(GadCurrent)
```

# Panic attacks ever
An abrupt surge of intense fear or intense discomfort that reaches a peak within minutes, and during which time four or more of a list of 13 physical and cognitive symptoms occur.

4 or more symptoms out of: 

Experienced the following symptom (SFP1): 
https://biobank.ndph.ox.ac.uk/ukb/coding.cgi?id=1910
13	Afraid I was going to die
12	Afraid of losing control or 'going crazy'
11	Felt like things weren't real or detached from self
10	Felt numbness or tingling sensations
9	Felt hot or cold
8	Felt dizzy, unsteady, light-headed or faint
7	Nauseous or felt sick in the stomach
6	Pain or discomfort in chest
5	Felt like I was choking
4	Felt short of breath, or like being smothered
3	Trembling or shaking 
2	Sweating
1	Heart was pounding or racing

0	No, I have not had this happen to me

-3	Prefer not to answer

4 or more panic attack symptoms must be endorsed

If participants endorse fewer than 4 panic symptoms, then they are given a 0 for diagnosis 
check colnames for field number 29065
29065-0.0 - 29065-0.12
```{r Check colnames}
anx_pad_data %>%
  select(contains("29065")) %>% 
  head()
# field name ranges: 29065-0.0 - 29065-0.12
```
Check summary and coding
```{r}
anx_pad_data %>%
  select(matches("^29065-0\\.\\d+_mod$")) %>% # matches regular expression (column names that start with "29065-0.", followed by a digit, and end with "_mod")
  summary()
```

Check frequency of panic attack symptom responses
```{r}
anx_pad_data %>%
  select(matches("^29065-0\\.\\d+_mod$")) %>% # matches regular expression (column names that start with "29065-0.", followed by a digit, and end with "_mod")
  freq()
```
Calculate panic attacks symptoms score          
```{r}
#Create score for panic attack symptoms
anx_pad_data <- anx_pad_data %>%
  mutate(PanicSymptomsCount = 0)

anx_pad_data <- anx_pad_data %>%
  mutate(PanicSymptomsCount = 
           case_when(
             `29065-0.0_mod` == 0 ~ 0,
             `29065-0.12_mod` > 0 ~ 13,
             `29065-0.11_mod` > 0 ~ 12,
             `29065-0.10_mod` > 0 ~ 11,
             `29065-0.9_mod` > 0 ~ 10,
             `29065-0.8_mod` > 0 ~ 9,
             `29065-0.7_mod` > 0 ~ 8,
             `29065-0.6_mod` > 0 ~ 7,
             `29065-0.5_mod` > 0 ~ 6,
             `29065-0.4_mod` > 0 ~ 5,
             `29065-0.3_mod` > 0 ~ 4,
             `29065-0.2_mod` > 0 ~ 3,
             `29065-0.1_mod` > 0 ~ 2,
             `29065-0.0_mod` > 0 ~ 1)
  )

anx_pad_data %>% 
  freq(PanicSymptomsCount)
```
Create panic attacks ever case 
```{r}
anx_pad_data <- anx_pad_data %>%
  mutate(PanicAttacksEver = case_when(is.na(PanicSymptomsCount) ~ NA,
                                    PanicSymptomsCount < 4 ~ 0,
                                     PanicSymptomsCount > 3 ~ 1))

anx_pad_data %>% 
  freq(PanicAttacksEver)
```
## Panic disorder
Symptoms {panic attack ever} 
AND 
Feel anxious more panic attacks (SFP3a) = Yes 
OR 
Feel worried bad things happening because of panic attacks (SFP3b) = Yes 
OR 
Avoid situations in which panic attacks might occur (SFP3c) = Yes 
AND 
Continue to worry about panic attacks or their consequences (SFP4) = Between 1 and 6 months (01) OR Between 6 and 12 months (02) OR Between 1 and 5 years (03) OR More than 5 years (04) OR All of my life / as long as I can remember (05) 
AND 
Ever the result of a medical condition, medication, drugs or alcohol (SFP7) = “No, never” (00) OR “Yes, some of them” (01) 
AND 
Do they occur in specific situations that cause you strong fear (SFP8) = No (00) OR “Yes, some of them” (01) 

Coding for 29070-0.0 (how long Continue to worry about panic attacks or their consequences)
0	Less than 1 month
1	Between 1 and 6 months
2	Between 6 and 12 months
3	Between 1 and 5 years
4	More than 5 years
5	All of my life / as long as I can remember
-1	Do not know
-3	Prefer not to answer

+++ ATK: Note to check with team on if we want to add more stringent control criteria (which is not specified in the MHQ2 algorithm guidance). i.e. dealing with those who are NA for some panic disorder diagnostic criteria, they should be kept as NA and not 0, or at the very least make everyone NA who did not respond to the MHQ2
```{r}
anx_pad_data <- anx_pad_data %>% 
  mutate(PanicDisorder = 
           case_when(PanicAttacksEver == 1 # Symptoms {panic attack ever}
                     & (`29067-0.0_mod` == 1 # Feel anxious more panic attacks = Yes
                     | `29068-0.0_mod` == 1 # OR Feel worried bad things happening because of panic attacks = Yes
                     | `29069-0.0_mod` == 1) # OR Avoid situations in which panic attacks might occur = Yes
                        & `29070-0.0_mod` > 0 # how long worry periods: codes 1-5 (greater than less than 1 month)
                     & (`29074-0.0_mod` == 0  # AND not due to other reasons 'no never'
                     | `29074-0.0_mod` == 1) # OR due to other reasons 'yes some of them'
                     & (`29075-0.0_mod`== 0 # occur in specific situation 'no'
                     | `29075-0.0_mod`== 1) # OR occur in specific situation 'Yes, some of them'
                     ~ 1, # code as 1 if meet all criteria
                     TRUE ~ 0 # other wise code as 0 
                       )
  )

anx_pad_data %>% 
  freq(PanicDisorder)
```
# Save the data
```{r save data}
anx_pad_data %>% 
  select(Gad7Score,
         GadCurrent,
         PanicSymptomsCount,
         PanicAttacksEver,
         PanicDisorder) %>% 
  saveRDS(
    file = "/file/path/anxiety_algorithms.rds")

```
