---
title: "UKB MHQ2 BED ICD-11"
author: "Helena Davies"
date: "2023-10-08"
output: html_document
---

```{r}
library("summarytools")
library("dplyr")
library("tidyverse")
```

```{r read in data}
dat <- readRDS("/Users/helenadavies/werk/UKB_MHQ2_algorithms/anon_data/MHQ2_Anonymous.rds")
```

*ICD-11 criteria*
1) Binge eating at least once a week over a period of at least three months. 
PLUS 
2) Either never weighed much less than other people expected 
OR 
Never had BMI indicating underweight 
OR 
Had periods when underweight, but binge eating never occurred when underweight 
PLUS 
3) At time when regular episodes of overeating, none of the listed ways of controlling weight / shape were used 


*Written algorithm*
Have you had recurrent episodes of excessive overeating or binge eating (EP6a) = 02 Yes, at least once a week 
AND 
What was the longest amount of time where you were overeating/binge eating at least once a week? (EP6b) = 03 At least three months 
 
AND 
NOT Case {Extended anorexia phenotype} 
OR 
BMI at low weight} > 18.55 kg/m2 
OR 
If you reported a time or times of low weight above, do/did you experience episodes of excessive overeating/binge eating during your time(s) of low weight? (EP6c) = 00 No, only at time(s) when I was NOT at low weight 
 
AND 
During your episodes of excessive overeating/binge eating, how often have you felt like you did not have control over your eating (e.g. not being able to stop eating or feeling compelled to eat)? (EP7) = 03 At least once a week for at least three months 
 
AND 
Do/did you feel distressed about your episodes of excessive overeating/binge eating? (EP9) =01 Yes 
 
AND 
During the time(s) when you were regularly overeating/binge eating, have you done any of the following as a way to control your body shape or weight? EP10= None of the above 

## Merge and freq

###29140 - Methods of controlling body shape or weight when overeating/binge eating

1              Made yourself vomit
2              Used laxatives (pills or liquids) [data 253 vs 500]
3              Used diuretics (water pills) [data 29 vs 145]
4              Used weight loss pills [191  vs 434]
5              Exercised excessively or felt distressed if unable to exercise [data 185 vs 441]
6              Fasted or not eaten for eight waking hours or more [data 285 vs 798]
7              Other methods to lose weight/stay at low weight [data 237 vs 622]
0              None of above
-3            Prefer not to answer

```{r}
setDT(dat,key = "ID")
dat[,`29140_n3`:=`29140-0.0`==-3 | `29140-0.1`==-3 | `29140-0.2`==-3 | `29140-0.3`==-3 | `29140-0.4`==-3 | `29140-0.5`==-3 | `29140-0.6`==-3]
freq(dat$`29140_n3`)
dat[,`29140_0`:=`29140-0.0`==0 | `29140-0.1`==0 | `29140-0.2`==0 | `29140-0.3`==0 | `29140-0.4`==0 | `29140-0.5`==0 | `29140-0.6`==0]
freq(dat$`29140_0`)
dat[,`29140_1`:=`29140-0.0`==1 | `29140-0.1`==1 | `29140-0.2`==1 | `29140-0.3`==1 | `29140-0.4`==1 | `29140-0.5`==1 | `29140-0.6`==1]
freq(dat$`29140_1`)
dat[,`29140_2`:=`29140-0.0`==2 | `29140-0.1`==2 | `29140-0.2`==2 | `29140-0.3`==2 | `29140-0.4`==2 | `29140-0.5`==2 | `29140-0.6`==2]
freq(dat$`29140_2`)
dat[,`29140_3`:=`29140-0.0`==3 | `29140-0.1`==3 | `29140-0.2`==3 | `29140-0.3`==3 | `29140-0.4`==3 | `29140-0.5`==3 | `29140-0.6`==3]
freq(dat$`29140_3`)
dat[,`29140_4`:=`29140-0.0`==4 | `29140-0.1`==4 | `29140-0.2`==4 | `29140-0.3`==4 | `29140-0.4`==4 | `29140-0.5`==4 | `29140-0.6`==4]
freq(dat$`29140_4`)
dat[,`29140_5`:=`29140-0.0`==5 | `29140-0.1`==5 | `29140-0.2`==5 | `29140-0.3`==5 | `29140-0.4`==5 | `29140-0.5`==5 | `29140-0.6`==5]
freq(dat$`29140_5`)
dat[,`29140_6`:=`29140-0.0`==6 | `29140-0.1`==6 | `29140-0.2`==6 | `29140-0.3`==6 | `29140-0.4`==6 | `29140-0.5`==6 | `29140-0.6`==6]
freq(dat$`29140_6`)
dat[,`29140_7`:=`29140-0.0`==7 | `29140-0.1`==7 | `29140-0.2`==7 | `29140-0.3`==7 | `29140-0.4`==7 | `29140-0.5`==7 | `29140-0.6`==7]
freq(dat$`29140_7`)
dat<-as.data.frame(dat)

```



## Not during AN criteria
From the ICD11: "The individual may be of normal weight or even low weight (though not sufficiently to meet the diagnostic requirements for Anorexia Nervosa)"
```{r not during AN criteria}
dat <- dat %>%
  mutate(not_during_AN =
           case_when(
             # "If you reported a time or times of low weight above, do/did you experience episodes of excessive overeating/binge eating during your time(s) of low weight?"
             
             # "Only at time(s) when I was not at low weight"
            (
             `29134-0.0` == 2 |
               
            # "Both at times of low weight and when not at low weight"   
             `29134-0.0` == 1 |
              
            # "Not applicable, as did not have low weight"
             `29134-0.0` == -4
            ) |
              
            # "Only at time(s) of low weight"
             (
               `29134-0.0` == 0 &
                
            # No lifetime anorexia nervosa
              DSM_AN_algorithm == 0 &
              self_reported_AN == 0 
            # +++ add any other reports of AN
            ) ~ 1
           )
  )
```

# Checks
For questions in algorithm
```{r sanity checks}
# Excessive overeating or binge eating
# Expecting 3,338 for "Yes, at least once a week" (2)
dat %>%
  freq(`29132-0.0`)

# Loss of control
# Expecting 3,904 for "At least once a week for at least 3 months" (0)
dat %>%
  freq(`29135-0.0`)

# Duration of overeating/binge eating
# Expecting 2,479 for "At least 3 months" (0)
dat %>%
  freq(`29133-0.0`)

# Distress re. binge eating
# Expecting 10,960 for "Yes" (1)
dat %>%
  freq(`29137-0.0`)
              
# Compensatory behaviours to control for body shape or weight
# Expecting 1,124 for "None of the above" (0)
dat %>%
  freq(`29140-0.0`)
                
# Binge eating during low weight
# Expecting 2,767 for "Only at time(s) when I was not at low weight" (2)
dat %>%
  freq(`29134-0.0`)
```

```{r BED ICD11 algorithm}
dat <- dat %>%
  mutate(ICD11_BED_algorithm =
           case_when(
             
        # "Have you had recurrent episodes of excessive overeating or binge eating (i.e. eating significantly more than what most people eat in a similar period of time, for example, two hours)?" = "Yes, at least once a week" 
            `29132-0.0` == 2 &
              
        # "During your episodes of excessive overeating/binge eating, how often have you felt like you did not have control over your eating (e.g. not being able to stop eating or feeling compelled to eat)?" = "At least once a week for at least 3 months"  
           `29135-0.0` == 0 &
              
        # "What was the longest amount of time when you were overeating/binge eating at least once a week?" = "At least 3 months"
           `29133-0.0` == 0 &
          
        # "Do/did you feel distressed about your episodes of excessive overeating/binge eating?" = "Yes"
          `29137-0.0` == 1 &
          
        # "During the time(s) when you were regularly overeating/binge eating, have you done any of the following as a way to control your body shape or weight? (Select all that apply)?" = "None of the above"
          `29140-0.0` == 0 &
        
       # ICD11 criteria: "The individual may be of normal weight or even low weight (though not sufficiently to meet the diagnostic requirements for Anorexia Nervosa)" (calculated above)
          not_during_AN == 1 ~ 1
           )
  )

# Check
dat %>%
  freq(ICD11_BED_algorithm)
```
