---
title: "UKB MHQ2 BED DSM-5"
author: "Helena Davies"
date: "2023-10-08"
output: html_document
---

```{r}
library("summarytools")
library("dplyr")
library("tidyverse")
```

```{r read in data}
dat <- readRDS("/Users/helenadavies/werk/UKB_MHQ2_algorithms/anon_data/MHQ2_Anonymous.rds")
```


*DSM-5 criteria*
1) Binge eating episodes at least once a week for three months. 
PLUS 
2) Either never weighed much less than other people expected OR Never had BMI indicating underweight OR 
Had periods when underweight, but binge eating never occurred when underweight 
PLUS 
3) Episodes feature loss of control at least once a week for at least three months 
PLUS 
4) Three or more of: 
Eat quickly 
Eat till uncomfortable 
Eat lots when not hungry 
Alone due to embarrassment 
Distress after eating 
PLUS 
5) Felt distressed about episodes of overeating 
PLUS 
6) Not associated with the recurrent use of inappropriate compensatory behaviour 


*Written algorithm*
Have you had recurrent episodes of excessive overeating or binge eating (EP6a) = 02 Yes, at least once a week 
AND 
What was the longest amount of time where you were overeating/binge eating at least once a week? (EP6b) = 03 At least three months 
 
AND 
NOT Case {Extended anorexia phenotype} 
OR 
BMI at low weight} > 18.55 kg/m2 
OR 
If you reported a time or times of low weight above, do/did you experience episodes of excessive overeating/binge eating during your time(s) of low weight? (EP6c) = 00 No, only at time(s) when I was NOT at low weight 
 
AND 
During your episodes of excessive overeating/binge eating, how often have you felt like you did not have control over your eating (e.g. not being able to stop eating or feeling compelled to eat)? (EP7) = 03 At least once a week for at least three months 
 
AND 
During these episodes of excessive overeating/binge eating, have you: 
(EP8) = At least 3 of: 
01 rapidly 
02 uncomfortably 
03 hungry 
04 embarassed 
05 disgusted 
 
AND 
Do/did you feel distressed about your episodes of excessive overeating/binge eating (EP9) =01 Yes 
 
AND 
During the time(s) when you were regularly overeating/binge eating, have you done any of the following as a way to control your body shape or weight? (EP10) = 00 None of the above 


# Create sumscore of binge eating behaviours (need at least 3 for DSM5 BED)
NB: Question was asked when Field 29132 ("Ever had recurrent episodes of excessive overeating or binge eating") was Yes.
```{r sumscore BE behaviours}
# Create numeric variables
dat <- dat %>%
  mutate(rapid_eating_numeric =
           case_when(`29136-0.0` == 1 |
                     `29136-0.1` == 1 |
                     `29136-0.2` == 1 | 
                     `29136-0.3` == 1 |
                     `29136-0.4` == 1 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(feeling_uncomf_full_numeric =
           case_when(`29136-0.0` == 2 |
                     `29136-0.1` == 2 |
                     `29136-0.2` == 2 | 
                     `29136-0.3` == 2 |
                     `29136-0.4` == 2 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(large_amounts_not_hungry_numeric =
           case_when(`29136-0.0` == 3 |
                     `29136-0.1` == 3 |
                     `29136-0.2` == 3 | 
                     `29136-0.3` == 3 |
                     `29136-0.4` == 3 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(eaten_alone_embarrassed_numeric =
           case_when(`29136-0.0` == 4 |
                     `29136-0.1` == 4 |
                     `29136-0.2` == 4 | 
                     `29136-0.3` == 4 |
                     `29136-0.4` == 4 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(disgusted_depressed_after_numeric =
           case_when(`29136-0.0` == 5 |
                     `29136-0.1` == 5 |
                     `29136-0.2` == 5 | 
                     `29136-0.3` == 5 |
                     `29136-0.4` == 5 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(no_BE_behaviours_numeric =
           case_when(`29136-0.0` == 0 |
                     `29136-0.1` == 0 |
                     `29136-0.2` == 0 | 
                     `29136-0.3` == 0 |
                     `29136-0.4` == 0 ~ 1,
                     
                     TRUE ~ 0
                     )
         ) %>%
  mutate(PNTA_BE_behaviours_numeric =
           case_when(`29136-0.0` == -3 |
                     `29136-0.1` == -3 |
                     `29136-0.2` == -3 | 
                     `29136-0.3` == -3 |
                     `29136-0.4` == -3 ~ 1,
                     
                     TRUE ~ 0
                     )
         )

# add statements!
# Check (expecting 6,312)
dat %>%
  freq(rapid_eating_numeric)

# Check (expecting 10,920)
dat %>%
  freq(feeling_uncomf_full_numeric)

# Check (expecting 9,963)
dat %>%
  freq(large_amounts_not_hungry_numeric)

# Check (expecting 5,798)
dat %>%
  freq(eaten_alone_embarrassed_numeric)

# Check (expecting 8,341)
dat %>%
  freq(disgusted_depressed_after_numeric)

# Check (expecting 2,429)
dat %>%
  freq(no_BE_behaviours_numeric)

# Check (expecting 97)
dat %>%
  freq(PNTA_BE_behaviours_numeric)


# Create sumscore
BE_behaviours_items <- dat %>% select(rapid_eating_numeric,
                                      feeling_uncomf_full_numeric,
                                      large_amounts_not_hungry_numeric,
                                      eaten_alone_embarrassed_numeric,
                                      disgusted_depressed_after_numeric)

dat <- dat %>%
    mutate(
        BE_behaviours_sumscore = rowSums(BE_behaviours_items,
                                         na.rm = TRUE)
    )

# Check (expecting maximum 5, minimum 0)
dat %>%
  descr(BE_behaviours_sumscore)
```
NB: This sumscore should only be used in the BED DSM5 algorithm, i.e., to indicate if participants have a score of 3 or above. A score of 0 is not informative because it includes people who did not answer the question or were not shown the question because of their answer to Field 29132.

# Checks
For remaining questions in algorithm
```{r sanity checks}
# Excessive overeating or binge eating
# Expecting 3,338 for "Yes, at least once a week" (2)
dat %>%
  freq(`29132-0.0`)

# Loss of control
# Expecting 3,904 for "At least once a week for at least 3 months" (0)
dat %>%
  freq(`29135-0.0`)

# Duration of overeating/binge eating
# Expecting 2,479 for "At least 3 months" (0)
dat %>%
  freq(`29133-0.0`)

# Distress re. binge eating
# Expecting 10,960 for "Yes" (1)
dat %>%
  freq(`29137-0.0`)
              
# Compensatory behaviours to control for body shape or weight
# Expecting 1,124 for "None of the above" (0)
dat %>%
  freq(`29140-0.0`)
                
# Binge eating during low weight
# Expecting 2,767 for "Only at time(s) when I was not at low weight" (2)
dat %>%
  freq(`29134-0.0`)
```

# BED DSM5 algorithm
```{r BED DSM5 algorithm}
dat <- dat %>%
  mutate(DSM5_BED_algorithm =
           case_when(
             
         # "Have you had recurrent episodes of excessive overeating or binge eating (i.e. eating significantly more than what most people eat in a similar period of time, for example, two hours)?" = "Yes, at least once a week" 
            `29132-0.0` == 2 &
        
        # "During your episodes of excessive overeating/binge eating, how often have you felt like you did not have control over your eating (e.g. not being able to stop eating or feeling compelled to eat)?" = "At least once a week for at least 3 months"  
           `29135-0.0` == 0 &
              
        # "What was the longest amount of time when you were overeating/binge eating at least once a week?" = "At least 3 months"
           `29133-0.0` == 0 &
          
        # "During these episodes of excessive overeating/binge eating, have you:" (Participant needs at least 3 of 5 binge-eating-related behaviours, sumscore calculated above)
          BE_behaviours_sumscore => 3 &
          
        # "Do/did you feel distressed about your episodes of excessive overeating/binge eating?" = "Yes"
          `29137-0.0` == 1 &
          
        # "During the time(s) when you were regularly overeating/binge eating, have you done any of the following as a way to control your body shape or weight? (Select all that apply)?" = "None of the above"
          `29140-0.0` == 0

          ( 
            # Extended anorexia phenotype != 1 | # HLD - why?
            # BMI at low weight > 18.55 kg/m2 | # HLD - why?

            # "If you reported a time or times of low weight above, do/did you experience episodes of excessive overeating/binge eating during your time(s) of low weight?" = "Only at time(s) when I was not at low weight" # HLD - shouldn't this also include "Both at times of low weight and when not at low weight" and "Not applicable, as did not have low weight"
            `29134-0.0`== 2 ~ 1)
           )
  )

# Check
dat %>%
  freq(DSM5_BED_algorithm)
```

