---
title: "MHQ2 Bulimia nervosa DSM-5 algorithm"
author: "Zhaoying YU, Helena Davies, Christopher Huebel, Johan Zvrskovec"
output: html_document
date: "2023-11-07"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, purl=FALSE}
library("tidyverse")
```

# Recode multiple choice questions
29140 Methods of controlling body shape or weight when overeating/binge eating
1	Made yourself vomit
2	Used laxatives (pills or liquids)
3	Used diuretics (water pills)
4	Used weight loss pills
5	Exercised excessively or felt distressed if unable to exercise
6	Fasted or not eaten for eight waking hours or more
7	Other methods to lose weight/stay at low weight
0	None of above
-3	Prefer not to answer

1	Made yourself vomit
```{r made yourself vomit}
dat <- dat %>%
  mutate(
    BE_made_yourself_vomit = case_when(
      `29140-0.0` == 1 |
        `29140-0.1` == 1 |
        `29140-0.2` == 1 |
        `29140-0.3` == 1 |
        `29140-0.4` == 1 |
        `29140-0.5` == 1 |
        `29140-0.6` == 1 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 581)
dat %>%
  count(BE_made_yourself_vomit)
```

2	Used laxatives (pills or liquids)
```{r used laxatives}
dat <- dat %>%
  mutate(
    BE_used_laxatives = case_when(
      `29140-0.0` == 2 |
        `29140-0.1` == 2 |
        `29140-0.2` == 2 |
        `29140-0.3` == 2 |
        `29140-0.4` == 2 |
        `29140-0.5` == 2 |
        `29140-0.6` == 2 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 500)
dat %>%
  count(BE_used_laxatives)
```

3	Used diuretics (water pills)
```{r used diuretics}
dat <- dat %>%
  mutate(
    BE_used_diuretics = case_when(
      `29140-0.0` == 3 |
        `29140-0.1` == 3 |
        `29140-0.2` == 3 |
        `29140-0.3` == 3 |
        `29140-0.4` == 3 |
        `29140-0.5` == 3 |
        `29140-0.6` == 3 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 145)
dat %>%
  count(BE_used_diuretics)
```

4	Used weight loss pills
```{r used weight loss pills}
dat <- dat %>%
  mutate(
    BE_used_weight_loss_pills = case_when(
      `29140-0.0` == 4 |
        `29140-0.1` == 4 |
        `29140-0.2` == 4 |
        `29140-0.3` == 4 |
        `29140-0.4` == 4 |
        `29140-0.5` == 4 |
        `29140-0.6` == 4 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 434)
dat %>%
  count(BE_used_weight_loss_pills)
```

5	Exercised excessively or felt distressed if unable to exercise
```{r excessive exercise}
dat <- dat %>%
  mutate(
    BE_excessive_exercise = case_when(
      `29140-0.0` == 5 |
        `29140-0.1` == 5 |
        `29140-0.2` == 5 |
        `29140-0.3` == 5 |
        `29140-0.4` == 5 |
        `29140-0.5` == 5 |
        `29140-0.6` == 5 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 441)
dat %>%
  count(BE_excessive_exercise)
```

6	Fasted or not eaten for eight waking hours or more
```{r fasten}
dat <- dat %>%
  mutate(
    BE_fasted = case_when(
      `29140-0.0` == 6 |
        `29140-0.1` == 6 |
        `29140-0.2` == 6 |
        `29140-0.3` == 6 |
        `29140-0.4` == 6 |
        `29140-0.5` == 6 |
        `29140-0.6` == 6 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 798)
dat %>%
  count(BE_fasted)
```

7	Other methods to lose weight/stay at low weight
```{r other weight control}
dat <- dat %>%
  mutate(
    BE_other_weight_control = case_when(
      `29140-0.0` == 7 |
        `29140-0.1` == 7 |
        `29140-0.2` == 7 |
        `29140-0.3` == 7 |
        `29140-0.4` == 7 |
        `29140-0.5` == 7 |
        `29140-0.6` == 7 ~ 1,
      # did not have any weight control behaviour
      `29140-0.0` == 0 ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )

# Check (expecting 622)
dat %>%
  count(BE_other_weight_control)
```

# Bulimia nervosa

```{r}
dat <- dat %>%
  mutate(
    Bulimia_nervosa = case_when(
      # Criterion A
      # Have you had recurrent episodes of excessive overeating
      # or binge eating = 2 Yes, at least once a week
      (
        `29132-0.0` == 2 &
          # AND During your episodes of excessive overeating/binge eating,
          # how often have you felt like you did not have control over your
          # eating = 0 At least once a week for at least three months
          # This is matches the Criterion C question
          `29135-0.0` == 0
      ) &
        # Criterion B
        # AND During the time(s) when you were regularly overeating or
        # binge eating, have you done any of the following
        # as a way to control your body shape or weight?
        # 1 Made yourself vomit OR
        # 2 Used laxatives (pills or liquids meant to
        # stimulate bowel movements) OR
        # 3 Used diuretics (water pills) OR
        # 4 Used weight loss pills (over the counter or prescription) OR
        # 5 Exercised excessively, felt compelled to exercise,
        # felt uneasy or distressed if unable to exercise or prioritised
        # exercise over your health or important activities OR
        # 6 Fasted or not eaten for eight waking hours or more
        (
          BE_made_yourself_vomit == 1 |
            BE_used_laxatives == 1 |
            BE_used_diuretics == 1 |
            BE_used_weight_loss_pills == 1 |
            BE_excessive_exercise == 1 |
            BE_fasted == 1
        ) &
        # Criterion C
        # AND During the time when you were regularly overeating/binge eating,
        # did you ever use any of these behaviours (made yourself vomit,
        # used pills, exercised excessively or fasted), on their own or
        # in combination, at least once a week? = 1 Yes, at least once a week
        (
          `29141-0.0` == 1 &
            # AND What was the longest amount of time where you were overeating
            # or binge eating at least once a week? = 0 At least three months
            `29133-0.0` == 0
        ) &
        # Criterion D
        # AND During the time when you were overeating/binge eating,
        # how dependent was your self-worth on your body shape or weight?
        # 2 A great deal OR 1 A moderate amount
        `29143-0.0` %in% c(1, 2) &
        # Criterion E
        # AND not during AN
        BE_not_during_AN == 1 ~ 1,
      # has answered ED section of MHQ2
      !is.na(`29205-0.0`) ~ 0,
      # people who have not answered the MHQ2
      TRUE ~ NA_real_
    )
  )
```


```{r, purl=FALSE}
dat %>%
  select(
    Bulimia_nervosa,
    `29132-0.0`, # binge eating
    `29135-0.0`, # frequency of loss of control
    BE_made_yourself_vomit,
    BE_used_laxatives,
    BE_used_diuretics,
    BE_used_weight_loss_pills,
    BE_excessive_exercise,
    BE_fasted,
    `29141-0.0`, # compensatory behaviours frequency
    `29133-0.0`, # duration binge eating
    `29143-0.0`, # self-esteem when binge eating
    BE_not_during_AN
  ) %>%
  filter(
    !is.na(`29132-0.0`)
  )
```

# Check
```{r, purl=FALSE}
dat %>%
  count(Bulimia_nervosa)
```
