---
title: "UKB MHQ2 Extended binge eating"
author: "Helena Davies"
date: "2023-10-08"
output: html_document
---

*Criteria*
Extended binge-eating phenotype 
Recurrent episodes of excessive overeating, regardless of length and context 

*Written algorithm*
Have you had recurrent episodes of excessive overeating or binge eating (EP6a) = 02 Yes, at least once a week 
Extended eating disorder phenotypes include participants that may not meet the exact diagnostic criteria used for disorder algorithms, and do not represent any clinical entity 

```{r}
library("summarytools")
library("dplyr")
library("tidyverse")
```

```{r read in data}
dat <- readRDS("/Users/helenadavies/werk/UKB_MHQ2_algorithms/anon_data/MHQ2_Anonymous.rds")
```

```{r extended BE}
dat <- dat %>%
  mutate(extended_BE_algorithm =
           case_when(
             
  # "Have you had recurrent episodes of excessive overeating or binge eating (i.e. eating significantly more than what most people eat in a similar period of time, for example, two hours)?" = "Yes, at least once a week" 
            `29132-0.0` == 2  ~ 1)
  )

# HLD - Shouldn't we add loss of control?
```

