---
title: "CIDI-SF Lifetime MDD MHQ2 UK Biobank"
author: "Danyang Li"
date: "09/10/2023"
output: html_document
---
Lifetime depression algorithm of MHQ2 UK Biobank 

Configuration of global options 
```{r Setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  prompt = FALSE,
  cache = FALSE
)
```

```{r Clean global environment, purl=FALSE}
rm(list = ls())
```

```{r Load packages, purl=FALSE}
library(data.table)
library(tidyverse)
```

```{r Read data, purl=FALSE}
setwd("/scratch/prj/ukbiobank/UKB_MHQ2_Anonymised")
dat <- readRDS("data/MHQ2_Anonymous.rds")
```

# 1. Depression ever case
Algorithm: 
Persistent sadness (29011) = Yes (1) OR Loss of interest (29012) = Yes (1)
AND
How much of day (29014) = Most of day (1) or All day long (0)  
AND
Did you feel this way (29015) = Almost every day (1) or Every day (0)
AND
Impairment (29031) = Somewhat (1) or A lot (0)
AND
Total number of symptoms endorsed (core and others) >= 5
Persistent sadness (core) 29011; 
Loss of interest (core) 29012; 
Tired or low energy 29018 (Yes (1)); 
Gain or loss of weight 29021 (gained weight (0) or lost weight (1) or gain and loss (2)); 
Sleep change 29022 (Yes (1)); 
Trouble concentrating 29026 (Yes (1)); 
Feeling worthless 29027 (Yes (1)); 
Thinking about death 29029 (Yes (1))
```{r Depression ever case}

if(diagnosticsFlag) unique(dat$`29014-0.0`)
if(diagnosticsFlag) unique(dat$`29021-0.0`)

cidid_symptoms <- c(
  "29011-0.0",
  "29012-0.0",
  "29018-0.0",
  "29022-0.0",
  "29026-0.0",
  "29027-0.0",
  "29029-0.0"
)

dat <-
  dat %>%
  mutate(
    # missingness of core symptoms (yes = either symptoms are missing)
    cidid.no.info = case_when(
      (
        is.na(`29011-0.0`) |
          `29011-0.0` < 0
      ) |
        (
          is.na(`29012-0.0`) |
            `29012-0.0` < 0
        ) ~ 1,
      .default = 0
    ),
    day.time = case_when(
      `29014-0.0` == 1 |
        `29014-0.0` == 0 ~ 1,
      `29014-0.0` == 2 |
        `29014-0.0` == 3 ~ 0,
      .default = NA_real_
    ),
    day.freq = case_when(
      `29015-0.0` == 1 |
        `29015-0.0` == 0 ~ 1,
      `29015-0.0` == 2 ~ 0,
      .default = NA_real_
    ),
    impairment = case_when(
      `29031-0.0` == 1 |
        `29031-0.0` == 0 ~ 1,
      `29031-0.0` == 2 |
        `29031-0.0` == 3 ~ 0,
      .default = NA_real_
    ),
    # modify the coding format of weight change
    `tmp_29021-0.0` = case_when(
      `29021-0.0` == 0 |
        `29021-0.0` == 1 |
        `29021-0.0` == 2 ~ 1,
      `29021-0.0` == 3 ~ 0,
      .default = NA_real_
    ),
    # change negative value to NA
    across(
      all_of(cidid_symptoms),
      ~ case_when(
        (. == 1) ~ 1,
        (. == 0) ~ 0,
        (
          (. < 0) |
            is.na(.)
        ) ~ NA_real_
      ),
      .names = "tmp_{col}"
    )
  ) %>%
  mutate(
    # core symptoms + frequency
    cidid.screen = case_when(
      (
        `29011-0.0` == 1 |
          `29012-0.0` == 1
      ) &
        day.time == 1 &
        day.freq == 1 &
        impairment == 1 ~ 1,
      (
        `29011-0.0` == 0 &
          `29012-0.0` == 0
      ) |
        day.time == 0 |
        day.freq == 0 |
        impairment == 0 ~ 0,
      .default = NA_real_
    )
  ) %>%
  mutate(
    # total number of symptoms
    cidid.symptom.score = rowSums(
      select(
        .,
        starts_with("tmp_")
      ),
      na.rm = TRUE
    ) * NA_real_^(
      rowSums(
        !is.na(
          select(
            .,
            starts_with("tmp_")
          )
        )
      ) == 0
    ) # anything ^ 0 equals 1, show NA if all items are missing
  ) %>%
  mutate(
    # depression ever case
    depression.ever.case = case_when(
      cidid.screen == 1 &
        !is.na(cidid.symptom.score) &
        cidid.symptom.score >= 5 ~ 1,
      (
        cidid.screen == 0 |
          cidid.symptom.score < 5
      ) &
        cidid.no.info == 0 &
        !is.na(cidid.symptom.score) ~ 0,
      .default = NA_real_
    )
  ) %>%
  select(
    -cidid.no.info,
    -cidid.screen,
    -cidid.symptom.score,
    -day.time,
    -day.freq,
    -impairment,
    -starts_with("tmp_")
  )

rm(cidid_symptoms)

```

# 2. Depression ever control
NOT mental health diagnosis (MHC1) (29000) = Depression (1)
AND NOT Persistent Sadness (29011) = Yes
AND NOT Loss of interest (29012) = Yes
AND NOT PHQ full score >= 5 (see below)
•	Little interest or pleasure (29002) = More than half of the days (2) or Nearly every day (3)
•	Feeling down (29003) = More than half of the days (2) or Nearly every day (3)
•	Trouble sleeping (29004) = More than half of the days (2) or Nearly every day (3)
•	Feeling tired (29005) = More than half of the days (2) or Nearly every day (3)
•	Poor appetite(29006)  = More than half of the days (2) or Nearly every day (3)
•	Feeling guilt (29007) = More than half of the days (2) or Nearly every day (3)
•	Trouble concentrating (29008) = More than half of the days (2) or Nearly every day (3)
•	Moving slowly (29009) = More than half of the days (2) or Nearly every day (3)
•	Thoughts death (29010) = Several days (1) More than half of the days (2) or Nearly every day (3)
```{r Depression ever control}

phq_symptoms <- c(
  "29002-0.0",
  "29003-0.0",
  "29004-0.0",
  "29005-0.0",
  "29006-0.0",
  "29007-0.0",
  "29008-0.0",
  "29009-0.0",
  "29010-0.0"
)

dat$tmp_29000 <- apply(
  dat[, grepl(
    "29000-0",
    colnames(dat)
  )],
  1,
  function(x) {
    "1" %in% x
  }
)

dat <-
  dat %>%
  # change coding format of PHQ 9 symptoms
  mutate(
    across(
      all_of(phq_symptoms), ~ case_when(
        (
          . == 2 |
            . == 3
        ) ~ 1,
        (
          . == 0 |
            . == 1
        ) ~ 0,
        (
          . < 0 |
            is.na(.)
        ) ~ NA_real_
      ), .names = "tmp_{col}"
    )
  ) %>%
  mutate(
    # PHQ score
    phq.symptom.score = rowSums(
      select(
        .,
        starts_with("tmp_")
      ),
      na.rm = TRUE
    ) * NA_real_^(
      rowSums(
        !is.na(
          select(
            .,
            starts_with("tmp_")
          )
        )
      ) == 0
    )
  ) %>%
  mutate(
    depression.ever.control = case_when(
      tmp_29000 == FALSE &
        `29011-0.0` == 0 &
        `29012-0.0` == 0 &
        phq.symptom.score < 5 &
        !is.na(phq.symptom.score) ~ 1,
      tmp_29000 == TRUE |
        `29011-0.0` == 1 |
        `29012-0.0` == 1 |
        (
          phq.symptom.score >= 5 &
            !is.na(phq.symptom.score)
        ) ~ 0,
      .default = NA_real_
    )
  ) %>%
  select(-starts_with("tmp_"))

rm(phq_symptoms)

```

# 3. Depression and subthreshold depressive symptoms ever
Persistent Sadness (29011) = Yes (1)
OR
D3 Loss of interest (29012) = Yes (1)
OR
mental health diagnosis (MHC1) (29000) = Depression (1)
OR
PHQ full score >= 5 (see below)
•	Little interest or pleasure (29002) = More than half of the days (2) or Nearly every day (3)
•	Feeling down (29003) = More than half of the days (2) or Nearly every day (3)
•	Trouble sleeping (29004) = More than half of the days (2) or Nearly every day (3)
•	Feeling tired (29005) = More than half of the days (2) or Nearly every day (3)
•	Poor appetite(29006)  = More than half of the days (2) or Nearly every day (3)
•	Feeling guilt (29007) = More than half of the days (2) or Nearly every day (3)
•	Trouble concentrating (29008) = More than half of the days (2) or Nearly every day (3)
•	Moving slowly (29009) = More than half of the days (2) or Nearly every day (3)
•	Thoughts death (29010) = Several days (1) More than half of the days (2) or Nearly every day (3)
```{r subthreshold depressive symptoms ever}

dat$tmp_29000 <- apply(
  dat[, grepl(
    "29000-0",
    colnames(dat)
  )],
  1,
  function(x) {
    "1" %in% x
  }
)

dat <-
  dat %>%
  mutate(
    subthreshold = case_when(
      (
        `29011-0.0` == 1 |
          `29012-0.0` == 1 |
          tmp_29000 == TRUE |
          (
            phq.symptom.score >= 5 &
              !is.na(phq.symptom.score)
          )
      ) ~ 1,
      `29011-0.0` == 0 &
        `29012-0.0` == 0 &
        tmp_29000 == FALSE &
        (
          phq.symptom.score < 5 &
            !is.na(phq.symptom.score)
        ) ~ 0,
      .default = NA_real_
    )
  ) %>%
  mutate(
    depression.subthreshold = case_when(
      depression.ever.case == 1 |
        subthreshold == 1 ~ 1,
      depression.ever.case == 0 &
        subthreshold == 0 ~ 0,
      .default = NA_real_
    )
  ) %>%
  select(
    -phq.symptom.score,
    -tmp_29000,
    -subthreshold
  )
```

+++JZ: Polar depresion and everything after point 3 moved to other script as these are dependent on mania
