---
title: 'CIDI-SF Lifetime MDD MHQ2 UK Biobank'
author: "Danyang Li"
date: "09/10/2023"
output: html_document
---
Lifetime depression algorithm of MHQ2 UK Biobank - 'polar' depression traits ONLY - these are dependent on the mania part

Configuration of global options 
```{r Setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = '',
  prompt = FALSE,
  cache = FALSE
  )
```

```{r Clean global environment, purl=FALSE}
rm(list = ls())
```

```{r Load packages, purl=FALSE}
library(data.table)
library(tidyverse)
```

```{r Read data, purl=FALSE}
setwd('/scratch/prj/ukbiobank/UKB_MHQ2_Anonymised')
# MHQ2 <- fread('ukb675453_MHQ2_Anonymised.csv', data.table=F)
MHQ2 <- readRDS('data/MHQ2_Anonymous.rds')
```

# 4. Single episode unipolar depression
Case {depression ever}
AND
Number of episodes (29033)= One (1)
AND
NOT case {bipolar type I}

Excluded if number of episodes missing or bipolar state missing 
```{r bipolar type I cases }
# this is the variable of biopolar type I from mania algorithm needed for single and recurrent unipolar depression

#bipolar.type.I 
# MHQ2$BD1: bipolar type 1 case = 1, not case = 0 #EXTERNAL FROM MANIA

```
```{r Single episode unipolar depression}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        depression.single = case_when(
            depression.ever.case == 1 & 
            `29033-0.0` == 1 & 
            BD1 == 0 ~ 1,
            depression.ever.case == 0 | 
            (`29033-0.0` > 1 & !is.na(`29033-0.0`)) | 
            `29033-0.0` == -4  | 
            BD1 == 1 ~ 0,
            .default = NA_real_
        )
    )

```

# 5. Recurrent unipolar depression 
Case {depression ever}
AND
Number of episodes (29033) = Several >1
AND
NOT case {bipolar type I}

Excluded if number of episodes missing or bipolar state missing
```{r Recurrent unipolar depression}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        depression.recurrent = case_when(
            depression.ever.case == 1 & 
            (!is.na(`29033-0.0`) & (`29033-0.0` > 1 | `29033-0.0` == -4)) & 
            BD1 == 0 ~ 1,
            depression.ever.case == 0 | 
            BD1 == 1 | 
            `29033-0.0` == 1  ~ 0,
            .default = NA_real_
        )
    )

```

# 6. Single episode unipolar depression triggered by event 
Case {depression single episode}
AND
start within two months of event (29013) = Yes (1)
```{r Single episode unipolar depression triggered by event}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        depression.single.event = case_when(
            depression.single == 1 & `29013-0.0` == 1 ~ 1,
            depression.single == 0 | `29013-0.0` == 0 ~ 0,
            .default = NA_real_
        )
    )
```

# 7. Post-natal depression 
Case {depression ever}
AND
Did this first episode occur within months of giving birth? Or has it been suggested that you had post-natal depression? (29035) = Yes (1)
```{r Post-natal depression}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        depression.postnatal = case_when(
            depression.ever.case == 1 & `29035-0.0` == 1 ~ 1,
            depression.ever.case == 0 | `29035-0.0` == 0 ~ 0,
            .default = NA_real_
        )
    )
```

# 8. Worst depressive episode likely melancholic features (melancholic-like depression) 
Case {depression ever}
AND
(Loss of interest (29012) = Yes (1)
OR
Did your mood brighten in response to positive events (29016) = No (0)
AND 
(> 2) of:
•	Was your mood worse (29017) = “in the morning” (0)
•	Appetite (29020) = “decreased appetite” (2) OR Weight change (29021) = “lost weight” (1)
(counts as 1 item)
•	Sleep change (29022) = Yes (1) AND Waking too early (29024) = Yes (1)
•	Guilt (29028) = Yes (1)

```{r Melancholic-like depression}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        mood.worse = case_when(
            `29017-0.0` == 0 ~ 1,
            `29017-0.0` == 1 | `29017-0.0` == 2 ~ 0,
            .default = NA_real_
        ),
        appetite.weight = case_when(
            (`29020-0.0` == 2 | `29021-0.0` == 1) ~ 1,
            (`29020-0.0` == 0 & (`29021-0.0` == 0 | `29021-0.0` == 2 | `29021-0.0` == 3)) | 
            (`29020-0.0` == 1 & (`29021-0.0` == 0 | `29021-0.0` == 2 | `29021-0.0` == 3)) ~ 0,
            .default = NA_real_
        ),
        wake.early = case_when(
            `29022-0.0` == 1 & `29024-0.0` == 1 ~ 1,
            `29022-0.0` == 0 | `29024-0.0` == 0 ~ 0,
            .default = NA_real_
        ),
        guilt = case_when(
            `29028-0.0` == 1 ~ 1,
            `29028-0.0` == 0 ~ 0,
            .default = NA_real_
        )
    ) %>% 
    mutate(
        melan.symptom.score = rowSums(select(., mood.worse, appetite.weight, wake.early, guilt), na.rm=TRUE) * NA_real_^(rowSums(!is.na(select(., mood.worse, appetite.weight, wake.early, guilt))) == 0)
    ) %>% 
    mutate(
        depression.melancholic = case_when(
            depression.ever.case == 1 & 
            (`29012-0.0` == 1 | `29016-0.0` == 0) &
            (melan.symptom.score > 2 & 
            !is.na(melan.symptom.score)) ~ 1,
            depression.ever.case == 0 | 
            (`29012-0.0` == 0 & `29016-0.0` == 1) 
            | melan.symptom.score <= 2 ~ 0,
            .default = NA_real_
        )
    ) %>% 
    select(-mood.worse, -appetite.weight, -wake.early, -guilt, -melan.symptom.score)

```

# 9. Worst depressive episode likely atypical features (atypical depression)
Case {depression ever}
AND
NOT Case {worst depressive episode melancholic}
AND
Did your mood brighten in response to positive events (29016) = Yes (1)
AND
(>1) of:
•	Appetite (29020) = “increased appetite” (1) OR Weight (29021) = “gained weight” (0)
(counts as 1 item)
•	Sleep change (29022) = Yes (1) AND Sleeping too much (29025) = Yes (1)
•	Heavy feelings (29019) = Yes (1)
•	Coping with rejection, even when not  depressed (29032) = “Yes, and this has caused problems in work or social relationships” (0)
```{r Atypical depression}
MHQ2 <- 
    MHQ2 %>% 
    mutate(
        appetite.weight = case_when(
            `29020-0.0` == 1 | `29021-0.0` == 0 ~ 1,
            (`29020-0.0` == 0 & (`29021-0.0` == 1 | `29021-0.0` == 2 | `29021-0.0` == 3)) | 
            (`29020-0.0` == 2 & (`29021-0.0` == 1 | `29021-0.0` == 2 | `29021-0.0` == 3)) ~ 0,
            .default = NA_real_
        ),
        sleep.much = case_when(
            `29022-0.0` == 1 & `29025-0.0` == 1 ~ 1,
            `29022-0.0` == 0 | `29025-0.0` == 0 ~ 0,
            .default = NA_real_
        ),
        heavy.feeling = case_when(
            `29019-0.0` == 1 ~ 1,
            `29019-0.0` == 0 ~ 0,
            .default = NA_real_
        ),
        rejection = case_when(
            `29032-0.0` == 0 ~ 1,
            `29032-0.0` == 1 | `29032-0.0` == 2 ~ 0,
            .default = NA_real_
        )
    ) %>%  
    mutate(
        atypical.symptom.score = rowSums(select(., appetite.weight, sleep.much, heavy.feeling, rejection), na.rm=TRUE) * NA_real_^(rowSums(!is.na(select(., appetite.weight, sleep.much, heavy.feeling, rejection))) == 0) 
    ) %>%   
    mutate(
        depression.atypical = case_when(
            depression.ever.case == 1 & 
            depression.melancholic == 0 &
            `29016-0.0` == 1 &
            (atypical.symptom.score > 1 & 
            !is.na(atypical.symptom.score)) ~ 1,
            depression.ever.case == 0 | 
            depression.melancholic == 1 | 
            `29016-0.0` == 0 | 
            atypical.symptom.score <= 1 ~ 0, 
            .default = NA_real_
        )
    ) %>% 
    select(-appetite.weight, -sleep.much, -heavy.feeling, -rejection, -atypical.symptom.score)
```

# extract the data
```{r extract data, purl=FALSE}
MHQ2 <- 
    MHQ2 %>% 
    select(depression.ever.case, depression.ever.control, depression.subthreshold, depression.single, depression.recurrent, depression.single.event, depression.postnatal, depression.melancholic, depression.atypical)

saveRDS(MHQ2, 'life_time_depression.rds')

```



