# Life Not Worth Living
Ever thought life was not worth living (29108) (HB1). Yes, once (01) OR Yes, more than once (02) (Symptoms)

```{r}
dat$Not.Worth.Living <- with(dat,
  case_when(
    is.na(`29108-0.0`) | `29108-0.0` < 0 ~ NA,
    TRUE ~ `29108-0.0`
  )
)
```

# Harm To Self Ever
Deliberately harmed yourself (29111) (HB3) Yes , once (01) OR Yes, more than once (02) (Case)

```{r}
dat$Self.Harm.Ever <- with(dat,
  case_when(
    is.na(`29111-0.0`) | `29111-0.0` < 0 ~ NA_real_,
    !is.na(`29111-0.0`) & `29111-0.0` > 0 ~ 1,
    .default = 0
  )
)
```

# Harm To Self Ever, 12m
Deliberately harmed yourself (29111) (HB3) = Yes, once (01) OR Yes, more than once (02) AND In the last 12 months (29114) (HB4) = Yes (01)

```{r}
dat$Self.Harm.Ever.12m <- with(dat,
  case_when(
    is.na(`29111-0.0`) | `29111-0.0` < 0 | is.na(`29114-0.0`) | `29114-0.0` < 0 ~ NA,
    !is.na(`29111-0.0`) & `29111-0.0` > 0 &
      !is.na(`29114-0.0`) & `29114-0.0` > 0 ~ 1,
    .default = 0
  )
)

```

# Suicide Attempt
Harmed self with intent (29116) (HB6) = Yes, once (01) OR Yes, more than once (02)

```{r}
dat$Suicide.Attempt <- with(dat,
  case_when(
    is.na(`29111-0.0`) | `29111-0.0` < 0 | is.na(`29116-0.0`) | is.na(`29116-0.0`) < 0 ~ NA,
    !is.na(`29111-0.0`) & `29111-0.0` > 0 &
      !is.na(`29116-0.0`) & `29116-0.0` > 0 ~ 1,
    .default = 0
  )
)

```

