---
title: "Social Situations"
author: Zhaoying YU
output: html_document
date: "2024-01-16"
---

## Set-up

```{r, purl=FALSE}
library("dplyr")
library("tidyverse")
```

## Import data
```{r, purl=FALSE}

dat <-
  readRDS("/Users/lilian/Desktop/dat_Anonymous.rds")

```

## Algorithms
# Social Isolation

If Household (SS1, 29162)= Only me (0). Score +1 
If See in person (SS2, 29163) = Never (0) or Once every few months (1). Score +1 
Attend weekly (SS4, 29167) = None of the above (0). Score +1 
EVALUATE Score > 1 = positive
 >>> I think this logic might be a bit complicated for the code, and since the criteria is if two out of the three is met, social isolation is 'positive'. So I wrote the code this way:
 
```{r}

dat <- dat %>%
  mutate(
    social_isolation = case_when(
      (
        is.na(`29162-0.0`) |
          `29162-0.0` < 0
      ) |
        (
          is.na(`29163-0.0`) |
            `29163-0.0` < 0
        ) |
        (
          is.na(`29167-0.0`) |
            `29167-0.0` < 0
        ) ~ NA_real_,
      (
        `29162-0.0` == 0 &
          `29163-0.0` %in% c(0, 1)
      ) |
        (
          `29162-0.0` == 0 &
            `29167-0.0` == 0
        ) |
        (
          `29163-0.0` %in% c(0, 1) &
            `29167-0.0` == 0
        ) ~ 1,
      TRUE ~ 0
    )
  )

```

# Virtually connected

Positive if: Video calls (SS2a, 29164) = About once a week (3), 2-4 times a week (4) or Daily or almost daily (5) 
OR Virtual weekly (SS4a, 29168) NOT= None of the above (0)

```{r}

dat <- dat %>%
  mutate(
    virtually_connected = case_when(
      (is.na(`29164-0.0`) |
         `29164-0.0` < 0) |
        (
          is.na(`29168-0.0`) |
            `29168-0.0` < 0
        ) ~ NA_real_,
      `29164-0.0` %in% c(3, 4, 5) |
        !`29168-0.0` == 0 ~ 1,
      TRUE ~ 0
    )
  )

```


# Short scale UCLA loneliness

Sum scores on individual items after converting to 1-3 (Hardly ever (1) to Often (3)): 
Companionship (SS7, 29172) , Left out (SS8, 29173) , Isolated (SS9, 29174)

There are two further questions included for compatibility / comparison with previous UKB surveys: SS3 “How often are you able to confide…” and SS6 “In tune with the people around you”. Those are not included in this score

Original coding:
0	Hardly ever
1	Some of the time
2	Often
-1	Do not know
-3	Prefer not to answer


```{r}

##Re-code scores to 1-3
#Companionship, 29172
dat <- dat %>%
  mutate(
    `29172-0.0_recoded_social` = recode(
      as.character(`29172-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "-1" = NA_real_,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )

#Left out, 29173
dat <- dat %>%
  mutate(
    `29173-0.0_recoded_social` = recode(
      as.character(`29173-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "-1" = NA_real_,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )

#Isolated, 29174
dat <- dat %>%
  mutate(
    `29174-0.0_recoded_social` = recode(
      as.character(`29174-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "-1" = NA_real_,
      "-3" = NA_real_,
      .default = NA_real_
    )
  )

##Sum of UCLA scores
dat$UCLA_loneliness_sum <- rowSums(
  dat[, c(
    "29172-0.0_recoded_social",
    "29173-0.0_recoded_social",
    "29174-0.0_recoded_social"
  )],
  na.rm = FALSE
)

dat <- dat[, !grepl(
  "_recoded_social$",
  colnames(dat)
)]

```


# Brief Resilience Scale

SUM
Bounce back (BRS1, 29175), recover (BRS3, 29177) and little trouble (BRS5, 29179), score:
1. Strongly disagree
2. Disagree
3. Neutral
4. Agree
5. Strongly agree

Hard time (BRS2, 29176), hard to snap back (BRS4, 29178), and setbacks (BRS6, 29180) score:
5. Strongly disagree
4. Disagree
3. Neutral
2. Agree
1. Strongly agree

DIVIDE BY 6

```{r}

##Re-code scores
#Bounce back, BRS1, 29175
dat <- dat %>%
  mutate(
    `29175-0.0_recoded_social` = recode(
      as.character(`29175-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "3" = 4,
      "4" = 5,
      .default = NA_real_
    )
  )

#Recover, BRS3, 29177
dat <- dat %>%
  mutate(
    `29177-0.0_recoded_social` = recode(
      as.character(`29177-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "3" = 4,
      "4" = 5,
      .default = NA_real_
    )
  )

#Little Trouble, BRS5, 29179
dat <- dat %>%
  mutate(
    `29179-0.0_recoded_social` = recode(
      as.character(`29179-0.0`),
      "0" = 1,
      "1" = 2,
      "2" = 3,
      "3" = 4,
      "4" = 5,
      .default = NA_real_
    )
  )

#Hard time, BRS2, 29176
dat <- dat %>%
  mutate(
    `29176-0.0_recoded_social` = recode(
      as.character(`29176-0.0`),
      "0" = 5,
      "1" = 4,
      "2" = 3,
      "3" = 2,
      "4" = 1,
      .default = NA_real_
    )
  )

#Hard to snap back, BRS4, 29178
dat <- dat %>%
  mutate(
    `29178-0.0_recoded_social` = recode(
      as.character(`29178-0.0`),
      "0" = 5,
      "1" = 4,
      "2" = 3,
      "3" = 2,
      "4" = 1,
      .default = NA_real_
    )
  )

#Setbacks, BRS6, 29180
dat <- dat %>%
  mutate(
    `29180-0.0_recoded_social` = recode(
      as.character(`29180-0.0`),
      "0" = 5,
      "1" = 4,
      "2" = 3,
      "3" = 2,
      "4" = 1,
      .default = NA_real_
    )
  )

##BRS score sum
dat <- dat %>%
  mutate(
    BRS = (
      `29175-0.0_recoded_social` +
        `29177-0.0_recoded_social` +
        `29179-0.0_recoded_social` +
        `29176-0.0_recoded_social` +
        `29178-0.0_recoded_social` +
        `29180-0.0_recoded_social`
    ) / 6
  )

dat <- dat[, !grepl(
  "_recoded_social$",
  colnames(dat)
)]

```
