§---
title: "Trauma"
author: "Abigail ter Kuile"
date: "2023-11-23"
output: html_document
---

# Set up
```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = '',
  prompt = FALSE,
  cache = FALSE
  )
```

Install packages
```{r Install packages, purl=FALSE}
#install.packages("tidyverse")
```

Load packages from library 
```{r Load packages from library, purl=FALSE}
library(tidyverse)
```

clear global environment
```{r Clear global environment, purl=FALSE}
rm(list=ls())
```

```{r Get system date, purl=FALSE}
date <- Sys.Date()
date
```

Read in data
```{r read in data file, purl=FALSE}
dat <- readRDS(file = "/file/path/MHQ2_Anonymous.rds")
```
 
# Make trauma field data frame (for easier inspecting)
Adverse life events in childhood and adulthood field names MHQ: 29076-29090 
https://biobank.ndph.ox.ac.uk/ukb/label.cgi?id=1505

# specify field numbers
```{r}
trauma_field_numbers <- as.character(29076:29090)
```

# select trauma field numbers into dataframe
```{r}
trauma_data <- dat %>% 
  select(matches(paste(trauma_field_numbers, 
                       collapse = "|")))

# Inspect data
# glimpse(trauma_data)
# 
# summary(trauma_data)
```


# Convert data to numeric and recode -3 (prefer not to answer) to NA
To avoid overwriting raw data, save changes in new columns with suffix "_mod" (modified)
```{r}
trauma_data <- trauma_data %>%
  mutate(across(everything(), ~ ., .names = "{.col}_mod")) %>%
  mutate_all(~ as.numeric(.)) %>%
  mutate_all(~ case_when(
    . == -3 ~ NA_real_,
    TRUE ~ .
  ))
```

# Inspect data of modified columns
```{r, purl=FALSE}
modified_columns <- trauma_data %>%
  select(ends_with("_mod"))

# Inspect modified columns
glimpse(modified_columns)
summary(modified_columns)
```

# Childhood adverse events exposure
Repeated from MHQ1
Based on answers to the five questions of Childhood Trauma Screen (CTS). Screen is positive if any item is positive.
CTS takes one question from each domain of the Childhood Trauma Questionnaire. Thresholds taken from thresholds for represented domains in the CTQ. Other scoring schemes are used for other surveys.


Felt loved (29076) = Sometimes true (02) OR Rarely true (01)
+++ ATK: why does the algorithm text not include 'never true' (0)? I have included 0 in the code below 
OR 
Physically abused by family (29077) = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04) 
OR 
Felt hated by family (29078) = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04) 
OR 
Sexually molested (29079) = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04) 
OR 
Someone to take to doctor when needed as a child (29080) = Often true (03), Sometimes true (02) OR Rarely true (01) 

Fields 29076-29080

Coding	Meaning
0	Never true
1	Rarely true
2	Sometimes true
3	Often true
4	Very often true
-3	Prefer not to answer 

++ATK how do we want to handle NAs? If we are screening positive if any item is positive, I don't think it makes sense to exclude/code as NA for all items if they are only NA for one item (as hashed out below)
```{r}
trauma_data <- trauma_data %>%
  mutate(
    TraumaChildhood = case_when(
       (is.na(`29076-0.0_mod`)
         & is.na(`29077-0.0_mod`)
         & is.na(`29078-0.0_mod`)
         & is.na(`29079-0.0_mod`)
         & is.na(`29080-0.0_mod`)) ~ NA_real_,
      (`29076-0.0_mod` < 3) # Felt loved = Never true (0) Rarely true (1) or Sometimes true (2) 
     | (`29077-0.0_mod` > 0) #Physically abused by family = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04)
     | (`29078-0.0_mod` > 0) # Felt hated by family = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04) 
     | (`29079-0.0_mod` > 0) # Sexually molested = Rarely true (01), Sometimes true (02), Often true (03) OR Very often true (04) 
     | (`29080-0.0_mod` < 4) # Someone to take to doctor when needed as a child = Often true (03), Sometimes true (02) OR Rarely true (01) OR Never (0)
    ~ 1,
    TRUE ~ 0   )
    )
  
trauma_data %>%
  with(table(TraumaChildhood, useNA="a"))
```

# Adult abuse events exposure
New (not in MHQ1)
Based on answers to questions about partner abuse. Screen positive if any item is positive
Positive responses may indicate intimate partner abuse. 

Stopped seeing friends (29081) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 

OR 

Belittled (29082) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 

OR 

Hurt (29083) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 

OR 

Touched without consent (29084) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 

OR 

Sex without consent (29085) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 

fields 29081-29085

Coding	Meaning
0	No, never
1	Yes, but not in the last 12 months
2	Yes, within the last 12 months
-3	Prefer not to answer
```{r}
trauma_data <- trauma_data %>%
  mutate(
    TraumaAdulthood = case_when(
      (is.na(`29081-0.0_mod`)
        & is.na(`29082-0.0_mod`)
        & is.na(`29083-0.0_mod`)
        & is.na(`29084-0.0_mod`)
        & is.na(`29085-0.0_mod`)) ~ NA_real_,
      (`29081-0.0_mod` > 0) # Stopped seeing friends (29081) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01)  
     | (`29082-0.0_mod` > 0) # Belittled (29082) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01)  
     | (`29083-0.0_mod` > 0) # Hurt (29083) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01)  
     | (`29084-0.0_mod` > 0) # Touched without consent (29084) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01) 
     | (`29085-0.0_mod` > 0) # Sex without consent (29085) = Yes, within the last 12 months (02) OR Yes, but not in the last 12 months (01)
    ~ 1,
    TRUE ~ 0   )
    )
  
trauma_data %>% 
  with(table(TraumaAdulthood, useNA="a"))
```

# Adverse events 12 months exposure
New
Based on answers to question about partner abuse and other adverse events. Score positive if any items positive.
Because this includes a wider range of events, it is not a subset of above “adult abuse events” 

Positive if any of: 

Stopped seeing friends (29081) = Yes, within the last 12 months (02) 

OR 

Belittled (29082) = Yes, within the last 12 months (02) 

OR 

Hurt (29083) = Yes, within the last 12 months (02)  

OR 

Touched without consent (29084) = Yes, within the last 12 months (02) 

OR 

Sex without consent (29085) = Yes, within the last 12 months (02) 

OR 

Assault (29086) = Yes, within the last 12 months (02) 

OR 

Injury or illness (29087) = Yes, within the last 12 months (02) 

OR 

Divorce (29088) = Yes, within the last 12 months (02) 

OR 

Death spouse (29089) = Yes, within the last 12 months (02) 

OR 

Death suicide (29090) = Yes, within the last 12 months (02) 

fields 29081-29090
Coding	Meaning
0	No, never
1	Yes, but not in the last 12 months
2	Yes, within the last 12 months
-3	Prefer not to answer
```{r}
trauma_data <- trauma_data %>%
  mutate(
    Trauma12Months = case_when(
      (is.na(`29081-0.0_mod`)
        & is.na(`29082-0.0_mod`)
        & is.na(`29083-0.0_mod`)
        & is.na(`29084-0.0_mod`)
        & is.na(`29085-0.0_mod`)
        & is.na(`29086-0.0_mod`)
        & is.na(`29087-0.0_mod`)
        & is.na(`29088-0.0_mod`)
        & is.na(`29089-0.0_mod`)
        & is.na(`29090-0.0_mod`)
      ) ~ NA_real_,
      (`29081-0.0_mod` == 2) # Stopped seeing friends (29081) = Yes, within the last 12 months (02)   
     | (`29082-0.0_mod` == 2) # Belittled (29082) = Yes, within the last 12 months (02) 
     | (`29083-0.0_mod` == 2) # Hurt (29083) = Yes, within the last 12 months (02)
     | (`29084-0.0_mod` == 2) # Touched without consent (29084) = Yes, within the last 12 months (02) 
     | (`29085-0.0_mod` == 2) # Sex without consent (29085) = Yes, within the last 12 months (02) 
     | (`29086-0.0_mod` == 2) # Assault (29086) = Yes, within the last 12 months (02)
     | (`29087-0.0_mod` == 2) # Injury or illness (29087) = Yes, within the last 12 months (02)
     | (`29088-0.0_mod` == 2) # Divorce (29088) = Yes, within the last 12 months (02) 
     | (`29089-0.0_mod` == 2) # Death spouse (29089) = Yes, within the last 12 months (02)
     | (`29090-0.0_mod` == 2) # Death suicide (29090) = Yes, within the last 12 months (02) 
    ~ 1,
    TRUE ~ 0   )
    )
  
trauma_data %>% 
  with(table(Trauma12Months, useNA="a"))
```

# Select columns for saving
```{r select columns}
selected_columns <- trauma_data %>% 
  select(TraumaChildhood,
         TraumaAdulthood,
         Trauma12Months)
```

# Add the selected columns back to the MHQ2 dataframe
```{r}
MHQ2_trauma <- cbind(dat,
              selected_columns)
```

# Perform garbage collection
```{r, purl=FALSE}
gc()
```

# Save the data
```{r save the data, purl=FALSE}
MHQ2_trauma %>% 
  saveRDS(
    file = "/file/path/trauma_algorithms.rds")
```




